BioTime

### 2022-04-27 analysis/01_filter_by_traits.Rmd

```{r}
pins::board_register_github(name = "github", repo = "karinorman/biodivTS_data", branch = "master")

biotime_data <- pins::pin_get("biotime-data", board = "github")
elton_mamm <- pins::pin_get("elton-mamm", board = "github")
elton_bird <- pins::pin_get("elton-bird", board = "github")
amphibio <- pins::pin_get("amphibio", board = "github")

@@

Final merged data set to save
```{r}
#save locally and in cache
usethis::use_data(bt_traitfiltered)
pins::pin(bt_traitfiltered, board = "github")
```

@@

#save locally and in cache
usethis::use_data(trait_ref)
pins::pin(trait_ref, board = "github")
```

Create table of info for each study
```{r}
metadata <- pins::pin_get("metadata", board = "github")

study_table <- biotime_traits %>%
  select(study_id, plot) %>%
  count(study_id, name = "n_plots") %>% 
  left_join(metadata %>% select(study_id, realm, climate, habitat, biome_map, taxa, 
                                organisms, start_year, end_year), by = "study_id") %>%
  left_join(coverage %>% select(study_id, coverage = perc), by = "study_id")

#save locally and in cache
usethis::use_data(study_table)
pins::pin(study_table, board = "github")
```

>>>>

```{r}
#load data, objects named the same as the filename

load(here::here("data/biotime_data.rda")) #load biotime_data
load(here::here("data/elton_mamm.rda")) #load elton_mamm
load(here::here("data/elton_bird.rda")) #load elton_bird
load(here::here("data/amphibio.rda")) #load amphibio

@@

Final merged data set to save
```{r}
#save locally and in cache
usethis::use_data(bt_traitfiltered)
```

@@

#save locally and in cache
usethis::use_data(trait_ref)
```

Create table of info for each study
```{r}
load(here::here("data/meta.rda")) #load meta

study_table <- biotime_traits %>%
  select(study_id, plot) %>%
  count(study_id, name = "n_plots") %>% 
  left_join(meta %>% select(study_id, realm, climate, habitat, biome_map, taxa, 
                                organisms, start_year, end_year), by = "study_id") %>%
  left_join(coverage %>% select(study_id, coverage = perc), by = "study_id")

#save locally and in cache
#usethis::use_data(study_table)
```

### 2022-04-27 analysis/03_rarefy_null_models.Rmd
Add description of script contents.

```{r}
library(tidyverse)
library(FD)
library(furrr)

pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master")
```

@@

Get species pools for each timeseries
```{r}
bt_traitfiltered <- pins::pin_get("bt-traitfiltered", board = "github")
trait_ref <- pins::pin_get("trait-ref", board = "github")

@@

#write out null table
usethis::use_data(null_table)
pins::pin(null_table, board = "github")
```

>>>>

This script calculates n null samples for each rarefied samples, calculates the metrics for those null samples, and outputs a table of the null model statistics.

```{r}
library(tidyverse)
library(FD)
library(furrr)
```

@@

Get species pools for each timeseries
```{r}
load(here::here("data/bt_traitfiltered.rda"))
load(here::here("data/trait_ref.rda"))

@@

#write out null table
usethis::use_data(null_table)
```

### 2022-04-27 analysis/05_model_metrics.Rmd

```{r}
library(tidyverse)
library(lme4)
library(broom.mixed)
library(lmerTest)
library(emmeans)
```
#### Read in data, get final dataset for models
Get data and put it in long form 
```{r}
pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master")

#get metadata
meta <- pins::pin_get("meta", board = "github") %>%
  select(study_id, realm, climate, habitat, protected_area, biome_map)

meta <- pins::pin_get("bt-traitfiltered", board = "github") %>%
  select(study_id, rarefyid, taxa) %>% distinct() %>%
  left_join(meta)

#get metric data
metrics <- pins::pin_get("rarefied-metrics", board = "github") %>%
  # pivot_longer(., cols = setdiff(everything(), one_of("rarefyID", "STUDY_ID", "SamplePool", "SampleN", "num_years", "duration", "startYear", "endYear", "YEAR", "cell", "rarefied", "type", "realm", "climate", "habitat", "biome_map", "taxa", "organisms", "cent_lat", "cent_long", "abundance_type", "BROAD_TYPE", "taxa_mod", "climate_mod")), names_to = "metric", values_to = "value") %>%
  rename_with(tolower, .cols = setdiff(everything(), one_of("rarefyID")))  %>%
  mutate(logvalue = log(value + 1)) %>% # create a column for log transformation that adds the absolute value integer closest to the most negative value to create all positive values
  left_join(meta, c("rarefyID" = "rarefyid"))

@@

pins::pin(model_data, board = "github")
```

@@

usethis::use_data(metric_model_table)
pins::pin(metric_model_table, board = "github")
```

@@

#get study level slopes for base, and levels for categorical random effect in non-base models
re_slopes <- pmap_dfr(model_df, get_study_slopes)
pins::pin(re_slopes, board = "github")

@@

#All model estimates in one dataframe
# slope_models <- bind_rows(duration_slope, startyear_slope)
# pins::pin(slope_models, board = "github")

>>>>

```{r}
library(tidyverse)
library(lme4)
library(broom.mixed)
library(lmerTest)
library(emmeans)
```
#### Read in data, get final dataset for models
Get data and put it in long form 
```{r}
#get data
load(here::here("data/meta.rda"))
load(here::here("data/bt_traitfiltered.rda"))
load(here::here("data/rarefied_metrics.rda"))

meta <- meta %>%
  select(study_id, realm, climate, habitat, protected_area, biome_map)

meta <- bt_traitfiltered %>% 
  select(study_id, rarefyid, taxa) %>% distinct() %>%
  left_join(meta)

rm(bt_traitfiltered)

#get metric data
metrics <- rarefied_metrics %>%
  rename_with(tolower, .cols = setdiff(everything(), one_of("rarefyID")))  %>%
  mutate(logvalue = log(value + 1)) %>% # create a column for log transformation that adds the absolute value integer closest to the most negative value to create all positive values
  left_join(meta, c("rarefyID" = "rarefyid"))

@@

usethis::use_data(model_data)
```

@@

usethis::use_data(metric_model_table)
```

@@

#get study level slopes for base, and levels for categorical random effect in non-base models
re_slopes <- pmap_dfr(model_df, get_study_slopes)

@@

#All model estimates in one dataframe
# slope_models <- bind_rows(duration_slope, startyear_slope)

### 2022-04-27 data-raw/01_elton.Rmd

```{r}
#save locally
usethis::use_data(elton_bird, elton_mamm)

#save to cache
pins::board_register_github(name = "github", repo = "karinorman/biodivTS_data", branch = "master")
pins::pin(elton_bird, board = "github")
pins::pin(elton_mamm, board = "github")
```

>>>>

```{r}
#save locally
usethis::use_data(elton_bird, elton_mamm)
```

### 2022-04-27 data-raw/02_biotime_ids.Rmd

### Or, from the grid processing script
```{r}
#file created by 01_biotime_grid.R
pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master", name = "github")
bt_grid <- pins::pin_get("bt-grid", board = "github")

data <- bt_grid %>%
  rename_with(tolower) %>%
  mutate(species = str_replace(species, "_", " "))

rm(bt_grid)
```

@@

```{r}
#save locally
usethis::use_data(biotime_data)

#save to cache
pins::pin(biotime_data, board = "github")
```

>>>>

### Or, from the grid processing script
```{r}
#file created by 01_biotime_grid.R
load(here::here("data/bt_grid_collate.rda"))

data <- bt_grid_collate %>%
  rename_with(tolower) %>%
  mutate(species = str_replace(species, "_", " "))

rm(bt_grid_collate)
```

@@

```{r}
#save locally
usethis::use_data(biotime_data)
```
